The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-09-07 talk/vmil2012/paper.tex
de-stress execution speed issues

\begin{abstract}
Tracing just-in-time (JIT) compilers record linear control flow paths,
inserting operations called guards at points of possible divergence. These
operations occur frequently in generated traces and therefore it is important to
design and implement them carefully to find the right trade-off between
execution speed, deoptimization,
and memory overhead. In this paper, we perform an empirical analysis of runtime
properties of guards. This is used to guide the design of guards in the RPython
tracing JIT.
% \o/
\end{abstract}

@@

Besides being common, guards have various costs associated with them.
Guards have a runtime cost, they take time to execute. Therefore it is
important to make the on-trace execution of guards as efficient as possible. On
the other hand, guards are possible deoptimization points. The recorded and
compiled path has to be left if a guard fails, returning control to the
interpreter. Therefore guards need enough associated information to enable
rebuilding the interpreter state. The memory overhead of this information
should be kept low. These constraints and trade-offs are what makes the design
and optimization of guards an important and non-trivial aspect of the construction
of a tracing just-in-time compiler.

%Section~\ref{sec:Evaluation} presents Figures about the absolute number of
%operations for each benchmark, and the overhead produced by the information
%stored at the different levels for the guards
In this paper we want to substantiate the aforementioned observations about guards and
describe based on them the reasoning behind their implementation in
RPython's tracing just-in-time compiler. The contributions of this paper are:
\begin{itemize}
  \item An analysis of guards in the context of RPython's tracing JIT,
  \item detailed measurements about the frequency and the
  overhead associated with guards, and
  \item a description about how guards are implemented in the high\-
  and low-level components of RPython's JIT and a description of the rationale behind the design.
\end{itemize}

>>>>

\begin{abstract}
Tracing just-in-time (JIT) compilers record linear control flow paths,
inserting operations called guards at points of possible divergence. These
operations occur frequently in generated traces and therefore it is important to
design and implement them carefully to find the right trade-off between
deoptimization, memory overhead, and (partly) execution speed.
In this paper, we perform an empirical analysis of runtime
properties of guards. This is used to guide the design of guards in the RPython
tracing JIT.
% \o/
\end{abstract}

@@

Besides being common, guards have various costs associated with them.
Guards are possible deoptimization points. The recorded and
compiled path has to be left if a guard fails, returning control to the
interpreter. Therefore guards need enough associated information to enable
rebuilding the interpreter state. The memory overhead of this information
should be kept low. On the other hand,
Guards have a runtime cost, they take time to execute. Therefore it is
important to make the on-trace execution of guards as efficient as possible.
These constraints and trade-offs are what makes the design
and optimization of guards an important and non-trivial aspect of the construction
of a tracing just-in-time compiler.

%Section~\ref{sec:Evaluation} presents Figures about the absolute number of
%operations for each benchmark, and the overhead produced by the information
%stored at the different levels for the guards
In this paper we want to substantiate the aforementioned observations about guards and
describe based on them the reasoning behind their implementation in
RPython's tracing just-in-time compiler. The contributions of this paper are:
\begin{itemize}
  \item An analysis of guards in the context of RPython's tracing JIT,
  \item detailed measurements about the frequency and the
  memory overhead associated with guards, and
  \item a description about how guards are implemented in the high\-
  and low-level components of RPython's JIT and a description of the rationale behind the design.
\end{itemize}
